---
title: "MBiol_Analysis_1064807"
output: html_document
date: "2025-04-07"
---

#Introduction
This project aims to explore the interface between religion and attitudes towards tigers in TATR, Maharashtra, India. The below code shows the analysis conducted for the four objectives of this study:
(1) To examine the influence of religious beliefs on tolerance toward tigers through Sacred Wildlife Value Orientations (WVOs) and attitudes toward tigers.
(2) To explore the impact of sociodemographic factors on the strength of religious beliefs. 
(3) To evaluate the applicability of a four-compartment WVO typology in TATR and its association with sociodemographic factors, attitudes, and tolerance toward tigers.
(4) To analyze the influence of belief in divine protection and WVO typology on risk perception of tigers. 

#Admin
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r Install and Load Packages}
#install.packages("dplyr")
library(dplyr)
#install.packages("psych")  
library(psych)
#install.packages("lavaan")  
library(lavaan)             
#install.packages("semPlot", type = "source")
library("semPlot")
#install.packages("ggmosaic")
#install.packages("ggplot2")
library("ggplot2")
library(ggpubr)
library(ggthemes)
library(cowplot)
library(robustlmm)
library(emmeans)
library(ggplot2)
```


```{r Load Data} 
#Loading in data
file_path <- "~/Desktop/University Work/Fourth Year/Coding and Computing/Questionnaire Results - Results_Sheet_Final.csv"
MBiol_data <- read.csv(file_path)
head(MBiol_data)
```

```{r Age Bracket Creation}

#People aged 35, roughly the median age, or below were classified as “younger”, whilst those older 35 were classified as “older” (“younger” ≤ 35< “older”)

    MBiol_data <- MBiol_data %>%
  mutate(
    Age_Bracket = case_when(
      Age <= 35 ~ "Younger",
      Age>35 ~ "Older"
    )
  )
    
head(MBiol_data)
```

#Internal Reliability Test using Cronbach's Alpha
```{r Religious Beleifs}
#Calcualting Cronbach's Alpha to measure the internal reliability of items for Religious Belief Strength
religious_beliefs <- MBiol_data %>%
  select(ReligiousBeliefs1, ReligiousBeliefs2, ReligiousBeliefs3, ReligiousBeliefs4)
alpha_religious <- psych::alpha(religious_beliefs)
(religious_beliefs)
print(alpha_religious)
```

```{r Sacred WVOs}
#Calcualting Cronbach's Alpha to measure the internal reliability of items for Sacred WVO
sacred_wvo <- MBiol_data %>%
  select(SacredWVO1, SacredWVO2, SacredWVO3)
alpha_sacred <- psych::alpha(sacred_wvo)
(sacred_wvo)
print(alpha_sacred)
```

```{r Attitudes}
#Calcualting Cronbach's Alpha to measure the internal reliability of items for Attitudes towards tigers
attitudes <- MBiol_data %>%
  select(Attitudes1, Attitudes2, Attitudes3, Attitudes4, Attitudes5)
alpha_attitudes <- psych::alpha(attitudes)
(attitudes)
print(alpha_attitudes)
```

```{r Mutualism Sub-Dimension Statements}
#Calcualting Cronbach's Alpha to measure the internal reliability of items for the two sub-dimensions for Mutualism
mutualism_alpha_pre <- MBiol_data %>%
  select(SocialAffiliation1, SocialAffiliation2, Caring1, Caring2)
alpha_mutualism <- psych::alpha(mutualism_alpha_pre)
(mutualism_alpha_pre)
print(alpha_mutualism)
```

```{r Domination Sub-Dimension Statements}
#Calcualting Cronbach's Alpha to measure the internal reliability of items for the two sub-dimensions for Domination
domination_alpha_pre <- MBiol_data %>%
  select(AppropriateUse1, AppropriateUse2, Hunting1, Hunting2)
alpha_domination <- psych::alpha(domination_alpha_pre)
(domination_alpha_pre)
print(alpha_domination)
```

```{r Active Risk Mitigation}
#Calcualting Cronbach's Alpha to measure the internal reliability of items for active risk mitigation
active_risk_mitigation <- MBiol_data %>%
  select(Avoidable_Risk, Risk_Reduction)
alpha_risk_mitigation <- psych::alpha(active_risk_mitigation)
(active_risk_mitigation)
print(alpha_risk_mitigation)
```

Based on the outputs, Sacred WVO3 was removed

```{r Updating the full MBiol_data dataset}

MBiol_data <- MBiol_data %>%
  select(-SacredWVO3)
MBiol_data
```


#Creating Composite Scores
Based on the outcomes of this Cronbach's alpha analysis, composite scores (means of the items) are calculated as measures for the latent constructs. 

```{r Creating composite scores}
# Composite score for Religious Beliefs
MBiol_data <- MBiol_data %>%
  mutate(ReligiousBeliefs_Score = rowMeans(select(., ReligiousBeliefs1, ReligiousBeliefs2, ReligiousBeliefs3), na.rm = TRUE))

# Composite score for Sacred WVO
MBiol_data <- MBiol_data %>%
  mutate(SacredWVO_Score = rowMeans(select(., SacredWVO1, SacredWVO2), na.rm = TRUE))

# Composite score for Attitudes
MBiol_data <- MBiol_data %>%
  mutate(Attitudes_Score = rowMeans(select(., Attitudes1, Attitudes3, Attitudes4), na.rm = TRUE))

# Composite score for Active Risk Mitigation
MBiol_data <- MBiol_data %>%
  mutate(Active_Risk_Mitigation_Score = rowMeans(select(., Avoidable_Risk, Risk_Reduction), na.rm = TRUE))

# View the updated dataset
head(MBiol_data)
```

#Classification Within the Four-Compartment Model
```{r Calculation of Mutualism and Domination Scores}
#Step 1: Calculate Mutualism and Domination Scores
# Define column names for Mutualism and Domination belief dimensions
mutualism_columns <- c("SocialAffiliation1", "SocialAffiliation2", "Caring1", "Caring2")
domination_columns <- c("Hunting1", "Hunting2", "AppropriateUse1", "AppropriateUse2")

# Calculate mean scores for each respondent
MBiol_data <- MBiol_data %>%
  mutate(
    Mutualism_Score = rowMeans(select(., all_of(mutualism_columns)), na.rm = TRUE),
    Domination_Score = rowMeans(select(., all_of(domination_columns)), na.rm = TRUE)
  )

head(MBiol_data)
```

```{r Classification of respondents}
# Step 2: Classify respondents into WVO Typologies
# Define scale midpoint for classification
scale_midpoint <- 2.5

# Add a new column for WVO Typology
MBiol_data <- MBiol_data %>%
  mutate(
    WVO_Typology = case_when(
      Domination_Score > scale_midpoint & Mutualism_Score <= scale_midpoint ~ "Traditionalist",
      Domination_Score <= scale_midpoint & Mutualism_Score > scale_midpoint ~ "Mutualist",
      Domination_Score > scale_midpoint & Mutualism_Score > scale_midpoint ~ "Pluralist",
      Domination_Score <= scale_midpoint & Mutualism_Score <= scale_midpoint ~ "Distanced"
    )
  )

head(MBiol_data)
```

#Classification of Views on Active Risk Mitigation (i.e. classified as either optimistic or pessimistic based on the composite score)

```{r Active Risk Perception Bracket Creation}

    MBiol_data <- MBiol_data %>%
  mutate(
    Optimist_Vs_Pessimist = case_when(
      Active_Risk_Mitigation_Score <= 2 ~ "Pessimist",
      Active_Risk_Mitigation_Score>2 ~ "Optimist"
    )
  )
    
head(MBiol_data)
```

#Structural Equation Modelling

Used to test the causal relationships, both direct and indirect, for this hypothesised pathway: Religious Beliefs → Sacred WVO → Attitudes Toward Tigers → Tolerance.

There are two main components to SEM: the measurement model and the structural model. The measurement model tests how well the latent constructs are measured by their respective observed indicators. Using Confirmatory Factor Analysis (CFA), the measurement model tests whether the observed variables load onto their respective latent constructs as hypothesized. The structural model then examines the relationships between latent variables. Model fit was assessed using multiple fit indices including: Root Mean Square Error of Approximation (RMSEA), Standardized Root Mean Square Residual (SRMR), Comparative Fit Index (CFI), and Non-Normed Fit Index (NNFI)/Tucker-Lewis Index (TLI).

Tolerance was rescaled from a 1-5 scale to a 1-4 scale so that it was compatible with the other constructs.
```{r Rescaling Tolerance}
# Convert ordered factor to numeric (ensuring proper conversion)
MBiol_data <- MBiol_data %>%
  mutate(Tolerance_Numeric = as.numeric(as.character(Tolerance))) 

# Rescale to range 1-4
MBiol_data <- MBiol_data %>%
  mutate(Tolerance_Rescaled = scales::rescale(Tolerance_Numeric, to = c(1, 4)))

# Check rescaled values
summary(MBiol_data$Tolerance_Rescaled)
```

```{r}
# Load package for CFA
library(lavaan)

# Define the CFA model based on EFA results
cfa_model <- '
  ReligiousBeliefs =~ ReligiousBeliefs1 + ReligiousBeliefs2 + ReligiousBeliefs3 
  SacredWVOs =~ SacredWVO1 + SacredWVO2 
  Attitudes =~ Attitudes1  + Attitudes3 + Attitudes4 
'
# Fit CFA Model
cfa_fit <- cfa(cfa_model, data = MBiol_data, estimator = "MLR") #MLR used rather than MLE because of non-normality and heteroscadisicity

# Display summary with fit indices
summary(cfa_fit, fit.measures = TRUE, standardized = TRUE)

# Get goodness-of-fit statistics
fitMeasures(cfa_fit, c("cfi", "tli", "rmsea", "srmr"))
```

```{r Fitting SEM - direct effects}
SEM_model_refined <- '
  # Direct Effects
  SacredWVO_Score ~ ReligiousBeliefs_Score
  Attitudes_Score ~ SacredWVO_Score
  Tolerance_Rescaled ~ Attitudes_Score
'

# Fit SEM Model using Maximum Likelihood (MLR for robustness)
fitSEM_refined <- sem(SEM_model_refined, 
              data = MBiol_data, 
              estimator = "MLR")

# Summary with standardized estimates
summary(fitSEM_refined, fit.measures = TRUE, standardized = TRUE)

# Parameter estimates
parameterEstimates(fitSEM_refined, standardized = TRUE)

# Assessing model fit
fitMeasures(fitSEM, c("cfi", "tli", "rmsea", "srmr"))
```

```{r Fitting SEM - indirect/mediation effects}
# Define indirect paths in lavaan
sem_model_refined_indirect <- '
  SacredWVO_Score ~ a * ReligiousBeliefs_Score
  Attitudes_Score ~ b * SacredWVO_Score
  Tolerance_Rescaled ~ c * Attitudes_Score

  # Indirect effects
  indirect1 := a * b   # Mediation of SacredWVO_Score
  indirect2 := a * b * c   # Full mediation to Tolerance
'

# Fit the model with bootstrapped standard errors
fit_indirect_refined <- sem(sem_model_refined_indirect, data = MBiol_data, se = "bootstrap", bootstrap = 5000)

# Summary with indirect effects
summary(fit_indirect_refined, standardized = TRUE, ci = TRUE)
```

```{r Making SEM figure}
semPaths(fit_indirect_refined, "std", edge.label.cex = 1.2, fade = FALSE)
```

#Sociodemographic Factors, WVO Typology, and Religious Beliefs 

I used robust linear mixed-effects models (RLMMs) to examine how religious belief scores varied by gender and age bracket. Initial attempts to model religious belief scores using generalized linear mixed models (GLMMs) with Gamma distributions revealed significant overdispersion and violations of model assumptions, as indicated by DHARMa residual diagnostics. This prompted the switch to RLMMs.

###Robust Linear Mixed-Effect Model for Religious Beleif Score and Sociodemographic Factors (Age and Gender)
```{r Fitting RLMM for RBS and sociodemographic factors}
# Fit the model
robust_model <- rlmer(
  ReligiousBeliefs_Score ~ Gender * Age_Bracket + (1 | Village),
  data = MBiol_data
)

# emmeans accepting rlmer object
class(robust_model) <- "lmerMod"

# Get estimated marginal means
emm <- emmeans(robust_model, ~ Gender * Age_Bracket)

# Convert to a data frame
emm_df <- as.data.frame(emm)

```

```{r Creating EMM Visualisation (Figure 12)}
# Create the plot
ggplot(emm_df, aes(x = Age_Bracket, y = emmean, color = Gender, group = Gender)) +
  geom_point(position = position_dodge(width = 0.3), size = 3) +
  geom_line(position = position_dodge(width = 0.3), linewidth = 0.8) +
  geom_errorbar(aes(ymin = emmean - SE, ymax = emmean + SE),
                position = position_dodge(width = 0.3), width = 0.2) +
    scale_color_manual(values = c("F" = "red", "M" = "orange")) +
  labs(
    y = "Estimated Religious Beliefs Score",
    x = "Age Bracket",
    title = "Estimated Marginal Means of Religious Beliefs Score",
    subtitle = "With 95% Confidence Intervals",
    color = "Gender"
  ) +
  geom_jitter(data = MBiol_data, aes(x = Age_Bracket, y = ReligiousBeliefs_Score, color = Gender), 
              alpha = 0.2, width = 0.1, height = 0) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold"),
    plot.subtitle = element_text(color = "gray40")
  )
```


A linear mixed-effects model was fitted to examine whether WVO typology predicts attitudes toward tigers, with village included as a random intercept to account for clustering at the village level. The Attitudes Score was log-transformed prior to analysis to address right-skew and improve model assumptions.
### Linear Mixed-Effect Model for Attitudes (log) and WVO Typology
```{r Fitting LMM for Attitudes and WVO typology}
# Step 1: Log-transform the outcome variable
MBiol_data$log_Attitudes_Score <- log(MBiol_data$Attitudes_Score)

# Step 2: Fit a linear mixed model with the transformed outcome
library(lme4)

log_attitudes_model <- lmer(
  log_Attitudes_Score ~ WVO_Typology + (1 | Village),
  data = MBiol_data
)

# Step 3: Check model summary
summary(log_attitudes_model)
```

### Linear Mixed-Effect Model for WVO typology and Religious Beliefs
```{r Fitting multinomial model for WVO typology and Religious Beliefs}
multinom_model <- multinom(WVO_Typology ~ ReligiousBeliefs_Score, data = MBiol_data)

# Create a sequence of ReligiousBeliefs_Score values for prediction
new_data <- data.frame(ReligiousBeliefs_Score = seq(min(MBiol_data$ReligiousBeliefs_Score),
                                                     max(MBiol_data$ReligiousBeliefs_Score),
                                                     length.out = 100))

# Predict probabilities
predicted_probs <- predict(multinom_model, newdata = new_data, type = "probs")


# Convert predictions into long format for ggplot
predicted_long <- cbind(new_data, as.data.frame(predicted_probs)) %>%
  pivot_longer(cols = -ReligiousBeliefs_Score, names_to = "WVO_Typology", values_to = "Probability")
```

```{r Plotting probabilities plot}
# Plot
ggplot(predicted_long, aes(x = ReligiousBeliefs_Score, y = Probability, color = WVO_Typology)) +
  geom_line(size = 1.2) +
  labs(title = "Predicted Probability of WVO Typology by Religious Belief Score",
       x = "Religious Belief Score",
       y = "Predicted Probability",
       color = "WVO Typology") +
  theme_minimal()
```


###Fisher's Exact Tests



```{r Fisher's Exact test for Gender vs WVO typology}
# Create a frequency table of Gender vs. WVO_Typology
gender_wvo_table <- table(MBiol_data$Gender, MBiol_data$WVO_Typology)

# View the contingency table
print(gender_wvo_table)

# Run the Chi-Square test
chi_test_gender_wvo <- chisq.test(gender_wvo_table, simulate.p.value = TRUE)
print(chi_test_gender_wvo)

fisher_gender_wvo <- fisher.test(gender_wvo_table)
print(fisher_gender_wvo)
```


```{r Fisher's Exact test for Age vs WVO typology}
# Create a frequency table of Gender vs. WVO_Typology
age_wvo_table <- table(MBiol_data$Age_Bracket, MBiol_data$WVO_Typology)
age_wvo_table


fisher_age_wvo <- fisher.test(age_wvo_table)
print(fisher_age_wvo)
```


```{r Fishers Exact test for WVO and Active Risk Mitigation}
# Create a contingency table
wvo_optimism_table <- table(MBiol_data$WVO_Typology, MBiol_data$Optimist_Vs_Pessimist)

# View the table (optional)
print(wvo_optimism_table)

# Perform the Fishers test

fishers_test_result_wvo_optimism_table <-fisher.test(wvo_optimism_table)

# View the results
print(fishers_test_result_wvo_optimism_table)

```

